Computation and comparison of the stable northeastern United States marine boundary layer

### 2020-12-07 Paper/methodology.tex
Adding first part of Mike Sprague's comments

To maintain consistency with the measured data, the turbulence
intensity (TI) is calculated using the TKE as
\begin{equation}
  \textrm{TI} =
  \frac{\sqrt{\frac{2}{3}\times\textrm{TKE}}}{\overline{U}_{horiz}} =
  \frac{\sqrt{\frac{1}{3}\left( \overline{u'u'} + \overline{v'v'} + \overline{w'w'}
      \right)}}{\overline{U}_{horiz}}
\end{equation}
The averaged wind speeds were enforced at the measurement height of
20 m, and the applied wind directions were consistent with the
predominant stable direction of 225 degrees southwest.

@@

\subsection{Computational methodology}
%Provide a general description of the LES codes that we use.

@@

\subsection{LES formulation}

@@

\begin{align}
  \frac{\partial}{\partial t}\left(\bar{\rho} k \right) + \frac{\partial}{\partial x_j} \bar{\rho} k \widetilde{u_j} &= \frac{\partial}{\partial x_j} \left (2 \nu_t \frac{\partial k}{\partial x_j} \right) - 2 \nu_t \widetilde{S_{ij}} \widetilde{S_{ij}} + \frac{g}{\theta_\circ} \tau_{\theta_i} - C_{\epsilon} \frac{k^{3/2}}{l} \label{eqn:ksgs-les}
\end{align}
where
\begin{align}
  C_\epsilon &= 0.93 & \nu_t &= C_\epsilon l \sqrt{k} \\
  \tau_{\theta i} &= - 2\ \frac{\nu_t}{\mathrm{Pr}_t}\ \frac{\partial \widetilde{\theta}}{\partial x_i} & \mathrm{Pr}_t &= \left( 1 + \frac{2l}{\Delta} \right)^{-1} \\
\end{align}
The choice of the length scale ($l$) depends on whether the stratification is positive
or negative, based on the recommendation by Moeng~\cite{Moeng1984}, and is given
by
\begin{align}
  l =
  \begin{cases}
    \Delta = \sqrt[3]{\Delta_x \Delta_y \Delta_z}, & \frac{\partial \widetilde{\theta}}{\partial z} < 0 \\
    0.76\, \sqrt{k}\, \frac{g}{\theta_\circ}\, \left( \frac{\partial \widetilde{\theta}}{\partial z}\right)^{-1/2}, &  \frac{\partial \widetilde{\theta}}{\partial z} > 0 \\
  \end{cases}
\end{align}

\subsubsection{\label{sec:wallmodelBC}Lower wall model BC}
The sub-filter scale stresses are specified at the terrain boundary following
Moeng~\cite{Moeng1984} as
\begin{align}
 (\tau_{xz})_0 &= \average{\tau_{xz}}_0 \frac{V_1 \average{\widetilde{u}_1} + \average{V_1} (\widetilde{u}_1 - \average{\widetilde{u}_1} ) }{ \average{V_1} \sqrt{\average{\widetilde{u}_1}^2 + \average{\widetilde{v}_1}^2 } } , \\
 (\tau_{yz})_0 &= \average{\tau_{yz}}_0 \frac{V_1 \average{\widetilde{v}_1} + \average{V_1} (\widetilde{v}_1 - \average{\widetilde{v}_1} ) }{ \average{V_1} \sqrt{\average{\widetilde{u}_1}^2 + \average{\widetilde{v}_1}^2 } } \\
  (\tau_{\theta z})_0 &= \average{\tau_{\theta z}}_0 \frac{V_1 ( \average{\widetilde{\theta}_1} - \theta_{\circ} ) + \average{V_1} (\widetilde{\theta}_1 - \average{\widetilde{\theta}_1} ) }{ \average{V_1} ( \average{\widetilde{\theta}_1} - \theta_{\circ} ) },
\end{align}
where all quantities with the subscript 1 are values at the first vertical grid
level, $V_1$ is the resolved wind speed at the first grid level and
$\average{\bullet}$ represents the horizontal planar average. The mean
sub-filter stresses are computed using the Monin-Obukhov similarity
law~\cite{Etling1996}. The formulation for the potential temperature flux could
potentially experience a singularity when the mean resolved potential
temperature at the first grid level $\average{\widetilde{\theta}_1}$ crosses the
reference temperature $\theta_{\circ}$. Hence the model for the sub-filter
potential temperature flux is reformulated using the Monin-Obukhov similarity
law as
\begin{equation}
(\tau_{\theta z})_0 = \left ( \frac{u_{\tau} \kappa}{ \phi_h} \right ) \frac{V_1 ( \average{\widetilde{\theta}_1} - \theta_{\circ} ) + \average{V_1} (\widetilde{\theta}_1 - \average{\widetilde{\theta}_1} ) }{ \average{V_1} },
\end{equation}
where $\phi_h$ is the non-dimensional potential temperature gradient from
Monin-Obukhov similarity law~\cite{Etling1996}.

@@

While the primary objective of Nalu-Wind within the ExaWind simulation suite is
to support blade-resolved simulations of wind turbines, it comes equipped
with all the physics models necessary for modeling atmospheric boundary flows.
In the present work, the prognostic equation for temperature is solved using the
enthalpy equation (Eq.~\ref{eqn:enthalpy-les}). The prognostic one-equation,
subgrid-scale kinetic energy equation (Eq.~\ref{eqn:ksgs-les}) is used for LES
turbulence closure. Two Picard iterations were employed within each time step to
minimize splitting errors and recover second-order accuracy in time.

@@

As part of a suite of physics codes within the open-source project ExaWind,
AMR-Wind enables more efficient and scalable simulations of wind power plants.
This is because it is built on top of the AMReX software framework, which
contains all of the functionality needed to develop massively parallel,
block-structured adaptive mesh refinement (AMR) applications \cite{AMReX_JOSS}.
Because AMR-Wind is limited to Cartesian block-structured grids, it is
specialized to handle ABL physics and wind turbine wakes.  The block-structured
grids enable more efficient algorithms such as Multi-Level Multi-Grid (MLMG)
\cite{AMReX_JOSS} and are well suited for next generation supercomputers that
use graphical processing units (GPUs).  In the current work, we focus on
applying AMR-Wind to ABL computations using statically refined, uniform meshes,
and will examine adaptive mesh capabilities in future studies.

AMR-Wind solves the incompressible Navier-Stokes equations and is generalized to
handle variable density and viscosity. Additionally, scalar transport equations
can be solved, such as potential temperature or turbulence models. The
discretization in AMR-Wind is based on the approximate projection method used in
IAMR \cite{almgren1998conservative} and incflo \cite{sverdrup2018highly}. It is
a semi-staggered scheme where the velocity and scalar variables are located at
cell centers and pressure is located at nodes. Pressure is also staggered in
time so that pressure and the pressure gradient (located at cell centers) are at
time $n+1/2$.  The time discretization is handled with a Crank Nicolson
approach where the right-hand side of the system of equations is evaluated at
time $n+1/2$. The advection term is handled explicitly using an upwind
finite-volume method. There are multiple options within AMR-Wind, but in this
paper only the high-order Godunov Piecewise Parabolic Method (PPM) is used
\cite{Colella1984}. The advection terms are projected to face centers at $n+1/2$
and are corrected using an exact MAC projection~\cite{almgren1998conservative},
which guarantees a divergence free flow. The diffusion terms can be handled
explicitly, semi-implicitly, or fully implicit and are discretized using a
second-order central difference formula.  For the simulations in this paper, we
use a fully implicit scheme, as the variable viscosity from the eddy viscosity
may cause time step restrictions. After the scalar equations and the momentum
equations are advanced in time, a nodal projection is used to approximately
correct the velocity field to make it divergence free.  The nodal projection is
based on a node centered finite element method \cite{almgren1998conservative}.

@@

The computational methodology for the AMR-Wind and Nalu-Wind LES codes
follows practices similar to the previous offshore ABL study \cite{cheung2020large}, with
some modifications to handle the stable stratification.  As in
\cite{cheung2020large}, the domain was a square prism geometry, with
the $x$ and $y$ coordinates aligned in the East and North directions,
respectively.  The domain size and mesh requirements were determined
through a numerical grid study.  As discussed in section
\ref{sec:gridstudy}, the horizontal dimensions of 750 m $\times$ 750 m
were found to be sufficient to capture any large scale structures in
the boundary layer, and in all cases the vertical dimension of 1000 m
was also used.  The grid study in section \ref{sec:gridstudy} also
showed that the stable stratification required greater mesh refinement
to capture the smaller turbulent scales, and a uniform cell size of
2.5m was adopted for the offshore LES computations across both codes.

@@

\subsubsection{Determination of surface roughness and prescribed temperature }
To match the measured TI conditions given in table
\ref{tab:CapeWindMeasurements}, the surface roughness and prescribed
surface temperature change were adjusted through an initial
trial-and-error process.  The final surface roughness of $z_0$=0.0005m
used for the stable offshore conditions (see Table~\ref{tab:z0tempparam})
matched the roughness used in the neutral 15m/s and unstable 5m/s and
10 m/s cases from the previous study \cite{cheung2020large}.  This
value of surface roughness is consistent with the measurements from
the North Sea \cite{taylor2001dependence}, which found values of $z_0$
ranging from $5 \times 10^{-5}$m to $5\times 10^{-3}$m.

>>>>

To maintain consistency with the measured data, the turbulence
intensity (TI) is calculated using the TKE as
\begin{equation}
  \textrm{TI} =
  \frac{\sqrt{\frac{2}{3}\times\textrm{TKE}}}{\overline{U}_{horiz}} =
  \frac{\sqrt{\frac{1}{3}\left( \overline{u'_1 u'_1} + \overline{u'_2 u'_2} + \overline{u'_3 u'_3}
      \right)}}{\overline{U}_{horiz}}
\end{equation}
The averaged wind speeds were enforced at the measurement height of
20 m, and the applied wind directions were consistent with the
predominant stable direction of 225 degrees southwest.

@@

\subsection{Computational methodology and LES formulation}
%Provide a general description of the LES codes that we use.

@@

% \subsection{LES formulation}

@@

\begin{align}
  \frac{\partial}{\partial t}\left(\bar{\rho} k \right) + \frac{\partial}{\partial x_j} \bar{\rho} k \widetilde{u_j} &= \frac{\partial}{\partial x_j} \left (2 \nu_t \frac{\partial k}{\partial x_j} \right) - 2 \nu_t \widetilde{S_{ij}} \widetilde{S_{ij}} + \frac{g}{\theta_\circ} \tau_{\theta_i} - C_{\epsilon} \frac{k^{3/2}}{l} \label{eqn:ksgs-les}
\end{align}
where
\begin{align}
  C_\epsilon &= 0.93 & \nu_t &= C_\epsilon l \sqrt{k} \\
  \tau_{\theta i} &= - 2\ \frac{\nu_t}{\mathrm{Pr}_t}\ \frac{\partial \widetilde{\theta}}{\partial x_i} & \mathrm{Pr}_t &= \left( 1 + \frac{2l}{\Delta} \right)^{-1} \\
\end{align}
The choice of the length scale ($l$) depends on whether the stratification is positive
or negative, based on the recommendation by Moeng~\cite{Moeng1984}, and is given
by
\begin{align}
  l =
  \begin{cases}
    \Delta = \sqrt[3]{\Delta_x \Delta_y \Delta_z}, & \frac{\partial \widetilde{\theta}}{\partial z} < 0 , \\
    0.76\, \sqrt{k}\, \frac{g}{\theta_\circ}\, \left( \frac{\partial \widetilde{\theta}}{\partial z}\right)^{-1/2}, &  \frac{\partial \widetilde{\theta}}{\partial z} > 0 .\\
  \end{cases}
\end{align}

\subsubsection{\label{sec:wallmodelBC}Lower wall model BC}
The sub-filter scale stresses are specified at the terrain boundary following
Moeng~\cite{Moeng1984} as
\begin{align}
 (\tau_{xz})_0 &= \average{\tau_{xz}}_0 \frac{V_1 \average{\widetilde{u}_1} + \average{V_1} (\widetilde{u}_1 - \average{\widetilde{u}_1} ) }{ \average{V_1} \sqrt{\average{\widetilde{u}_1}^2 + \average{\widetilde{v}_1}^2 } } , \\
 (\tau_{yz})_0 &= \average{\tau_{yz}}_0 \frac{V_1 \average{\widetilde{v}_1} + \average{V_1} (\widetilde{v}_1 - \average{\widetilde{v}_1} ) }{ \average{V_1} \sqrt{\average{\widetilde{u}_1}^2 + \average{\widetilde{v}_1}^2 } } , \\
  (\tau_{\theta z})_0 &= \average{\tau_{\theta z}}_0 \frac{V_1 ( \average{\widetilde{\theta}_1} - \theta_{\circ} ) + \average{V_1} (\widetilde{\theta}_1 - \average{\widetilde{\theta}_1} ) }{ \average{V_1} ( \average{\widetilde{\theta}_1} - \theta_{\circ} ) },
\end{align}
where all quantities with the subscript 1 are values at the first vertical grid
level, $V_1$ is the resolved wind speed at the first grid level and
$\average{\bullet}$ represents the horizontal plane average at that height. The mean
sub-filter stresses are computed using the Monin-Obukhov similarity
law~\cite{Etling1996}. The formulation for the potential temperature flux could
potentially experience a singularity when the mean resolved potential
temperature at the first grid level $\average{\widetilde{\theta}_1}$ crosses the
reference temperature $\theta_{\circ}$. Hence the model for the sub-filter
potential temperature flux is reformulated using the Monin-Obukhov similarity
law as
\begin{equation}
(\tau_{\theta z})_0 = \left ( \frac{u_{\tau} \kappa}{ \phi_h} \right ) \frac{V_1 ( \average{\widetilde{\theta}_1} - \theta_{\circ} ) + \average{V_1} (\widetilde{\theta}_1 - \average{\widetilde{\theta}_1} ) }{ \average{V_1} },
\end{equation}
where $\phi_h$ is the non-dimensional potential temperature gradient from
Monin-Obukhov similarity law~\cite{Etling1996}.

@@

While the primary objective of Nalu-Wind within the ExaWind simulation suite is
to support blade-resolved simulations of wind turbines, it comes equipped
with all the physics models necessary for modeling atmospheric boundary flows.
In the present work, the prognostic equation for temperature is solved using the
enthalpy equation (Eq.~\ref{eqn:enthalpy-les}). The prognostic one-equation,
subgrid-scale kinetic energy equation (Eq.~\ref{eqn:ksgs-les}) is used for LES
turbulence closure. Two Picard iterations were employed within each time step to
reduce splitting errors and improving time accuracy.

@@

As part of a suite of physics codes within the open-source ExaWind simulation environment,
AMR-Wind enables more efficient and scalable simulations of wind power plants.
This is because it is built on top of the AMReX software framework, which
contains all of the functionality needed to develop massively parallel,
block-structured adaptive mesh refinement (AMR) applications \cite{AMReX_JOSS}.
Because AMR-Wind is limited to Cartesian block-structured grids, it is
specialized to handle ABL physics and wind turbine wakes.  The block-structured
grids enable more efficient algorithms such as Multi-Level Multi-Grid (MLMG)
\cite{AMReX_JOSS} and are well suited for next generation supercomputers that
use graphical processing units (GPUs).  In the current work, we focus on
applying AMR-Wind to ABL computations using statically refined, uniform meshes,
and will examine adaptive mesh capabilities in future studies.

AMR-Wind solves the incompressible Navier-Stokes equations and is generalized to
handle variable density and viscosity. Additionally, scalar transport equations
can be solved, such as potential temperature or turbulence models. The
discretization in AMR-Wind is based on the approximate projection method used in
IAMR \cite{almgren1998conservative} and incflo \cite{sverdrup2018highly}. It is
a semi-staggered scheme where the velocity and scalar variables are located at
cell centers and pressure is located at nodes. Pressure is also staggered in
time so that pressure and the pressure gradient (located at cell centers) are at
time $n+1/2$.  The time discretization is handled with a Crank Nicolson
approach where the right-hand side of the system of equations is evaluated at
time $n+1/2$. The advection term is handled explicitly using an upwind
finite-volume method. There are multiple options within AMR-Wind, but in this
paper only the high-order Godunov Piecewise Parabolic Method (PPM) is used
\cite{Colella1984}. The advection terms are projected to face centers at $n+1/2$
and are corrected using an exact MAC projection~\cite{almgren1998conservative},
which guarantees a divergence-free flow. The diffusion terms can be handled
explicitly, semi-implicitly, or implicitly and are spatially discretized using a
second-order central difference formula.  For the simulations in this paper, we
use a fully implicit scheme, as the variable viscosity from the eddy viscosity
may cause time step restrictions. After the scalar equations and the momentum
equations are advanced in time, a nodal projection is used to approximately
correct the velocity field to make it divergence free.  The nodal projection is
based on a node centered finite element method \cite{almgren1998conservative}.

@@

The computational methodology for the AMR-Wind and Nalu-Wind LES codes
follows practices similar to the previous offshore ABL study \cite{cheung2020large}, with
some modifications to handle the stable stratification.  As in
\cite{cheung2020large}, the domain was a square prism geometry, with
the $x$ and $y$ coordinates aligned in the East and North directions,
respectively.  The domain size and mesh requirements were determined
through a numerical grid study.  As discussed in section
\ref{sec:gridstudy}, the horizontal dimensions of 750 m $\times$ 750 m
were found to be sufficient to capture any large scale structures in
the boundary layer, and in all cases the vertical dimension of 1000 m
was also used.  The grid study in section \ref{sec:gridstudy} also
showed that the stable stratification required greater mesh refinement
to capture the smaller turbulent scales, and a uniform cell size of
2.5 m was adopted for the offshore LES computations across both codes.

@@

\subsubsection{Determination of surface roughness and prescribed temperature }
To match the measured TI conditions given in Table
\ref{tab:CapeWindMeasurements}, the surface roughness and prescribed
surface temperature change were adjusted through an initial
trial-and-error process.  The final surface roughness of $z_0$=0.0005 m
used for the stable offshore conditions (see Table~\ref{tab:z0tempparam})
matched the roughness used in the neutral 15 m/s and unstable 5 m/s and
10 m/s cases from the previous study \cite{cheung2020large}.  This
value of surface roughness is consistent with the measurements from
the North Sea \cite{taylor2001dependence}, which found values of $z_0$
ranging from $5 \times 10^{-5}$m to $5\times 10^{-3}$ m.
